Combinatorial Auctions

### 2015-12-17 paper.tex
afd

Combinatorial auctions are used to sell multiple goods at the same time and are particularly relevant when selling complementary and substitutive goods. In many ways combinatorial auctions are very similar to the standard auctions and matching markets that were covered in EECS 498.3 but the interdependencies between goods add aditional layers of complexity that make them very interesting to analyze. In the following paper, I will summerize and analyze part of Blumrosen and Nisan's survey on combinatorial auctions focusing on applications of the combinatorial auction, the single minded bidder, the walrasian equilibrium, and bidding languages \cite{paper}.

@@

In a standard auction, there is a single good being sold to one or more players who have a demand for that good. The auctioneer determines which player is allocated the good based on which player displays the greatest demand. There are many ways for the auctioneer to coordinate the auction, such as the ascending and descending price auctions that we discussed in class. Combinatorial auctions generalize this concept of an auction by allowing multiple goods to be sold at once to players that have demands for bundles or subsets of goods. The authors, Liad Blumrosen and Noam Nisan, in their paper on Combinatorial Auctions, analyze the computational complexity of possible allocation algorithms, how to represent and communicate the value functions of bidders, and the strategic behavior of rational bidders in combinatorial auctions. In the remainder of this paper, I will be summarizing and analyzing the main arguments made by Blumrosen and Nisan.

@@

As in any auction, the auctioneer in a combinatorial auction has to determine an allocation of goods. In combinatorial auctions the auctioneer has the add constraint that any one good can only be allocated to one player. Due to this constraint every allocation in a combinatorial auction is in the form $S_1,...,S_n$ where $S_i \cap S_j = \emptyset$ for every $i \neq j$. And from this, the social welfare of the allocation is equal to $\sum_i v_i(S_i)$.

In the remainding sections of this paper, I will discuss some real world applications of combinatorial auctions, a simplified bidder (the single minded bidder) that the authors introduce to simplify the complexity and strategic analysis of the auction, what is a walrasian equilibrium, how bidders can communicate their valuations through bidding languages, and finally, how we can operate a combinatorial ascending price auction.

@@

When the London bus system was deregulated in 1984, the commitee in charge needed to determine the best way to commission companies to operate  different bus routs across London. Ultimately the commitee in charge decided to use a form of a combinatorial first price auction to commission the bus routes. They made this decision so companies could take advantage of economies of scale by bidding on multiple bus routes at a time. Not only could companies bid on multiple routes at a time but by using a combinatorial auction the commission allowed companies to bid on bundles of routes without requiring the company to also bid on the individual routes seperatly. This allowed companies to take advantage of the benifits of owning similar bus routes, such as lower fixed costs, without the risk of being allocated only some of the desired similar bus routes. By using a combinatorial auction, the commission made it possible for companies to more accuratly express their valuations than if they had auctioned off each route seperatly, this turn lead to a more socially optimal allocation of the bus routes.

@@

There are three charicteristics of combinatorial auctions that the authors primarily analyze: computational complexity, strategic behavior, and representing the valuation functions. To simplify the analysis of computational complexity and strategic behavior, the authors remove the issue of representing the valuation functions by introducing the single minded bidder who has a very simple valuation function. The single minded bidder has a valuation function $v$ such that for a set $S'$ and value $v'$, $v(S) = v'$ for all $S \supseteq S'$ and $v(S) = 0$ for all other $S$. This simplification allows for a valuation function that is very easy to represent, namely the pair $(S', v')$, and therefore simplifies the complexity and strategy analysis.

\subsection{Complexity Analysis} % % % % % % % % % % % % 
An allocation for the combinatorial auction, as with all auctins, strives to maximize social welfare, which is defined as $\sum_i v_i(S_i)$. Since an allocation must allocate disjoint sets of goods to each winning bidder, to maximize social welfare, we should either allocate a bundle equal to a bidders demand bundle $S'$ or allocate the $\emptyset$. By doing so, there are no goods that are being wasted by being allocated to a bidder that doesn't gain value for them. Therefore the authors define an allocation to single minded bidders as follows:

@@

\begin{description}
  \item [Approximation:] Not only is it known that the independent set problem is NP-complete but it is also known that approximation the independent set problem is NP-Hard. Therefore, approximatin the optimal allocation of a combinatorial auction is NP-Hard.

@@

In the course, we found that the dominant strategy for bidders in a second-price auction is to bid their true value. This notition of bidding your ture value is called incentive compatibility. Incentive compatibility is beneficial so that the auctioneer (and consequently the allocation algorithm) knows that it is working with true information. The other option is for bidders to shade their bid, which adds another layer of the complexity of determining how bidders are expected to shade their bid to the allocation algorithm. Since there are many benifits of having an incentive compatible auction, the author's asked will the single minded bidder bid truthfully in a combinatorial auction?

@@

While the authors note that VCG and even VCG like prices cannot be used to produce an incentive compatibility auction because VCG only is appropriate for exactly optimal situations, I would argue that the construction of the prices for the greedy algorithm described by the authors is very much in the spirit if VCG prices. VCG prices are based on the ``harm'' that a bidder causes to all of the other bidders by participating in the auction. So in a standard second priced auction the top bidder pays a price equal to the bid of the second highest bidder because if the top bidder had not participate then the second highest bidder would have won the auction. Similarly in the greedy algorithm, a player is charged the price that is proportional to the bid of the player that was not allocated any good beacuse of the original player. While it is not as obvious with sets of goods as compared to single good, the prices that are charged in the greedy algorithm can be thought as the harm that the player is causing by participating in the auction.

@@

Using this definition of a demand of a bidder, the authors define a walrasian equilibrium as a set of ``market clearing'' prices where every bidder recieves a bundle in his demand set, and unallocated items have prices of zero. This result also shows that a Walrasian Equilibrium will be socially optimal, this is known as The First Welfare Theorem.

% \begin{displayquote}
As an example, there are two players Jack and Jill and two items \{x, y\}. Jack has the following value-item pairs \{3, x\} and \{6, x and y\} and Jill has the following valuations \{4, y\} and \{6, x and y\}. A walrasian equilibrium exists at prices 3 for $x$ and 4 for $y$. At these prices, Jack will win good $x$ and Jill will win good $y$ which gives a total social welfare of $3 + 4 = 7$ which is the maximum that is possible for this auction. To prove that these prices are in equilibrium, we can deviate the prices and see that the overall social welfare decreases. If the price of good $x$ goes down to 2 then the demand of both Jack and Jill becomes \{x and y\} and since both Jack and Jill cannot both get x and y then either Jack or Jill will not be allocated their demand. Either Jack or Jill will be allocated their demand making the social welfare equal to 6 which is less than the maximum of 7. If the price of $x$ goes up to 4 then Jack then no one will demand good $x$ and $x$ will be unallocated but since it has a price that is non-zero this violates the definition of a Walrasian equilibrium. Similarly, the price of $y$ can be fluxuated and shown to also decrease the social welfare of the allocation. Since changing both the prices of $x$ and $y$ are shown to decrease the total social welfare, it is shown that the prices are a walrasian equilibrium.
% \end{displayquote}

@@

Up until this point we have assumed that the auctioneer is just given the valuation function, $v_i$ of every bidder. In reality, for $m$ number of goods each bidder has to communicate their value for the $2^m - 1$ number of non-empty bundles of goods. Sinse this process becomes cobersome after only just a few dozen items, this communication becomes the limiting factor in our algorithm because we are able to compute allocations efficiently for thousands of goods. For this reason, it is important to discuss different mechanisms for communicating valuation functions to the auctioneer.

@@

By using a combination of ORing and XORing atomic bids it is possible to create very complex valuation functions that accuratly describe a players valuations.

@@

Being able to represent any valuation function solely as a ORing of atomic bids is very powerful for the following reason. A bidder that submits many ORed atomic bids looks the same as multiple bidders that each submitted a single atomic bid to an allocation algorithm. This is because the allocation algorithm is just trying to optimize the total social welfare so the algorithm doesn't care who is getting the value from the set that is being allocated but just cares about how that set is effecting the total social welfare. This means that all of the analysis and algorithms that were discussed for single minded bidders can be applied to valuation functions of ORs and XORs, which are much more expressive and accurate than the simplified signle minded bids.

@@

Combinatorial aucitions can be applied to many applications where a set of goods needs to be auctioned off and when players have valuations that depend on bundles of goods and not just single goods, such as the London bus routes, airport runway slots, and FCC spectrum auctions. Blumerosen and Nisan outlined a survey of many theoretical consideratoins when analyzing and designing a combinatorial auction and I summerized and analyzed a few of these topics \cite{paper}. We started with the single minded bidder, whose valuation pair $(S',v')$ provided a simple way to begin analyzing the computational complextity and strategic behaiver of the auction. We next discussed walrasian equilibria, which possess market clearing prices where every player is allocated a bundle in their demand. And finally, we discussed bidding languages and how bidders can effectivly communicate their valuations to the auctioneer. While bidding langagues turned out to be the computational bottle neck of the algorithm, we still determined a way to represent very complex valuations and while still being able to apply all of the analysis and algorithms of single minded bidders.

One of the key skills that was taught in the course was the ability to simplify a general definition of a concept based on specific constraints for a particular problem. For example, finding the page rank of an undirected and connect graph. I think that it would be benificial to develop a set of simpilar, special cases of combinatorial auctions that can then be applied alone or in combination of different real world applications. By analyzing more special cases, it is then more likely that a new real world application can be applied to a special case instead of the general auction which would make running and analyzing the auction simpilar.

>>>>

Combinatorial auctions are used to sell multiple goods at the same time and are particularly relevant when selling complementary and substitutive goods. In many ways combinatorial auctions are very similar to the standard auctions and matching markets that were covered in EECS 498.3 but the interdependencies between goods add additional layers of complexity that make them very interesting to analyze. In the following paper, I will summarize and analyze part of Blumrosen and Nisan's survey on combinatorial auctions focusing on applications of the combinatorial auction, the single minded bidder, the walrasian equilibrium, and bidding languages \cite{paper}.

@@

In a standard auction, there is a single good being sold to one or more players who have a demand for that good. The auctioneer determines which player is allocated the good based on which player displays the greatest demand. There are many ways for the auctioneer to coordinate the auction, such as the ascending and descending price auctions that we discussed in class. Combinatorial auctions generalize this concept of an auction by allowing multiple goods to be sold at once to players that have demands for bundles or subsets of goods. The authors, Liad Blumrosen and Noam Nisan, in their paper on Combinatorial Auctions, analyze the computational complexity of possible allocation algorithms, how to represent and communicate the value functions of bidders, and the strategic behavior of rational bidders in combinatorial auctions. In the remainder of this paper, I will be summarizing and analyzing the main arguments made by Blumrosen and Nisan \cite{paper}.

@@

As in any auction, the auctioneer in a combinatorial auction has to determine an allocation of goods. In combinatorial auctions the auctioneer has the added constraint that any one good can only be allocated to one player. Due to this constraint every allocation in a combinatorial auction is in the form $S_1,...,S_n$ where $S_i \cap S_j = \emptyset$ for every $i \neq j$. And from this, the social welfare of the allocation is equal to $\sum_i v_i(S_i)$.

In the remaining sections of this paper, I will discuss some real world applications of combinatorial auctions, a simplified bidder (the single minded bidder) that the authors introduce to simplify the complexity and strategic analysis of the auction, what is a walrasian equilibrium, and finally, how bidders can communicate their valuations through bidding languages.

@@

When the London bus system was deregulated in 1984, the committee in charge needed to determine the best way to commission companies to operate  different bus routs across London. Ultimately the committee in charge decided to use a form of a combinatorial first price auction to commission the bus routes. They made this decision so companies could take advantage of economies of scale by bidding on multiple bus routes at a time. Not only could companies bid on multiple routes at a time but by using a combinatorial auction the commission allowed companies to bid on bundles of routes without requiring the company to also bid on the individual routes separately. This allowed companies to take advantage of the benefits of owning similar bus routes, such as lower fixed costs, without the risk of being allocated only some of the desired similar bus routes. By using a combinatorial auction, the commission made it possible for companies to more accurately express their valuations than if they had auctioned off each route separately, this turn lead to a more socially optimal allocation of the bus routes.

@@

There are three characteristics of combinatorial auctions that the authors primarily analyze: computational complexity, strategic behavior, and representing the valuation functions. To simplify the analysis of computational complexity and strategic behavior, the authors remove the issue of representing the valuation functions by introducing the single minded bidder who has a very simple valuation function. The single minded bidder has a valuation function $v$ such that for a set $S'$ and value $v'$, $v(S) = v'$ for all $S \supseteq S'$ and $v(S) = 0$ for all other $S$. This simplification allows for a valuation function that is very easy to represent, namely the pair $(S', v')$, and therefore simplifies the complexity and strategy analysis.

\subsection{Complexity Analysis} % % % % % % % % % % % % 
An allocation for the combinatorial auction, as with all auctions, strives to maximize social welfare, which is defined as $\sum_i v_i(S_i)$. Since an allocation must allocate disjoint sets of goods to each winning bidder, to maximize social welfare, we should either allocate a bundle equal to a bidders demand bundle $S'$ or allocate the $\emptyset$. By doing so, there are no goods that are being wasted by being allocated to a bidder that doesn't gain value for them. Therefore the authors define an allocation to single minded bidders as follows:

@@

\begin{description}
  \item [Approximation:] Not only is it known that the independent set problem is NP-complete but it is also known that approximation the independent set problem is NP-Hard. Therefore, approximating the optimal allocation of a combinatorial auction is NP-Hard.

@@

In the course, we found that the dominant strategy for bidders in a second-price auction is to bid their true value. This notion of bidding your true value is called incentive compatibility. Incentive compatibility is beneficial so that the auctioneer (and consequently the allocation algorithm) knows that it is working with true information. The other option is for bidders to shade their bid, which adds another layer of the complexity of determining how bidders are expected to shade their bid to the allocation algorithm. Since there are many benefits of having an incentive compatible auction, the author's asked will the single minded bidder bid truthfully in a combinatorial auction?

@@

While the authors note that VCG and even ``VCG like'' prices cannot be used to produce an incentive compatibility auction because VCG only is appropriate for exactly optimal situations, I would argue that the construction of the prices for the greedy algorithm described by the authors is very much in the spirit if VCG prices. VCG prices are based on the ``harm'' that a bidder causes to all of the other bidders by participating in the auction. So in a standard second priced auction the top bidder pays a price equal to the bid of the second highest bidder because if the top bidder had not participate then the second highest bidder would have won the auction. Similarly in the greedy algorithm, a player is charged the price that is proportional to the bid of the player that was not allocated any good because of the original player. While it is not as obvious with sets of goods as compared to single good, the prices that are charged in the greedy algorithm can be thought of as the harm that the player is causing by participating in the auction.

@@

Using this definition of a demand of a bidder, the authors define a walrasian equilibrium as a set of ``market clearing'' prices where every bidder receives a bundle in his demand set, and unallocated items have prices of zero. This result also shows that a Walrasian Equilibrium will be socially optimal, this is known as The First Welfare Theorem.

% \begin{displayquote}
As an example, there are two players Jack and Jill and two items \{x, y\}. Jack has the following value-item pairs \{3, x\} and \{6, x and y\} and Jill has the following valuations \{4, y\} and \{6, x and y\}. A walrasian equilibrium exists at prices 3 for $x$ and 4 for $y$. At these prices, Jack will win good $x$ and Jill will win good $y$ which gives a total social welfare of $3 + 4 = 7$ which is the maximum that is possible for this auction. To prove that these prices are in equilibrium, we can deviate the prices and see that the overall social welfare decreases. If the price of good $x$ goes down to 2 then the demand of both Jack and Jill becomes \{x and y\} and since both Jack and Jill cannot both get x and y then either Jack or Jill will not be allocated their demand. Either Jack or Jill will be allocated their demand making the social welfare equal to 6 which is less than the maximum of 7. If the price of $x$ goes up to 4 then no one will demand good $x$ and $x$ will be unallocated but since it has a price that is non-zero this violates the definition of a Walrasian equilibrium. Similarly, the price of $y$ can be fluctuated and shown to also decrease the social welfare of the allocation. Since changing both the prices of $x$ and $y$ are shown to decrease the total social welfare, it is shown that the prices are a walrasian equilibrium.
% \end{displayquote}

@@

Up until this point we have assumed that the auctioneer is just given the valuation function, $v_i$ of every bidder. In reality, for $m$ number of goods each bidder has to communicate their value for the $2^m - 1$ number of non-empty bundles of goods. Since this process becomes cumbersome after only just a few dozen items, this communication becomes the limiting factor in our algorithm because we are able to compute allocations efficiently for thousands of goods. For this reason, it is important to discuss different mechanisms for communicating valuation functions to the auctioneer.

@@

By using a combination of ORing and XORing atomic bids it is possible to create very complex valuation functions that accurately describe a players valuations.

@@

Being able to represent any valuation function solely as a ORing of atomic bids is very powerful for the following reason. A bidder that submits many ORed atomic bids looks the same as multiple bidders that each submitted a single atomic bid to an allocation algorithm. This is because the allocation algorithm is just trying to optimize the total social welfare so the algorithm doesn't care who is getting the value from the set that is being allocated but just cares about how that set is effecting the total social welfare. This means that all of the analysis and algorithms that were discussed for single minded bidders can be applied to valuation functions of ORs and XORs, which are much more expressive and accurate than the simplified single minded bids.

@@

Combinatorial auctions can be applied to many applications where a set of goods needs to be auctioned off and when players have valuations that depend on bundles of goods and not just single goods, such as the London bus routes, airport runway slots, and FCC spectrum auctions. Blumrosen and Nisan outlined a survey of many theoretical considerations when analyzing and designing a combinatorial auction and in this paper I summarized and analyzed a few of these topics \cite{paper}. We started with the single minded bidder, whose valuation pair $(S',v')$ provided a simple way to begin analyzing the computational complexity and strategic behaver of the auction. We next discussed walrasian equilibria, which possess market clearing prices where every player is allocated a bundle in their demand. And finally, we discussed bidding languages and how bidders can effectively communicate their valuations to the auctioneer. While bidding languages turned out to be the computational bottle neck of the auction algorithm, we still determined a way to represent very complex valuations while still being able to apply all of the analysis and algorithms for single minded bidders.

One of the key skills that was taught in the course was the ability to simplify a general definition of a concept based on specific constraints for a particular problem. For example, finding the page rank of an undirected and connect graph. I think that it would be beneficial to develop a set of simpler, special cases of combinatorial auctions that can then be applied alone or in combination to different real world applications. By analyzing more special cases, it is then more likely that a new real world application can be applied to a special case instead of the general auction which would make running and analyzing the auction simpler.
